---
title: "malaria_atlas_project"
author: "John Archer"
date: "Jan - March 2021"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


Install/load required packages.
```{r packages}

# install.packages("gdistance")
# install.packages("abind")
# install.packages("rje")
# install.packages("ggplot2")
# install.packages("malariaAtlas")
# install.packages("sp")

library(gdistance)
library(abind)
library(rje)
library(ggplot2)
library(malariaAtlas)
library(sp)

## Plot defaults
theme_set(theme_minimal(base_size=14))

```


Create shapefile based on extent of 'leastcost_motor'.
```{r create_shapefile}

extent(leastcost_motor) 
# class      : Extent 
# xmin       : 33.19985 
# xmax       : 33.80001 
# ymin       : -11.29021 
# ymax       : -10.72993 

# Create bounding box
coords <- matrix(c(33.19985, -10.72993,
                   33.19985, -11.29021,
                   33.80001, -11.29021,
                   33.80001, -10.72993),
                 ncol = 2,
                 byrow = TRUE)

# Create polygon
aoi_polygon <- Polygon(coords)
aoi_polygon <- SpatialPolygons(list(Polygons(list(aoi_polygon), ID = "a")), proj4string=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

# Check
plot(aoi_polygon, axes = TRUE)

```


Check list of avaliable raster datasets
```{r}

# Generate and check list
# available_rasters <- listRaster(printed = TRUE)
View(available_rasters)

```
Raster data of interest:
"Global friction surface enumerating land-based travel speed with access to motorized transport for a nominal year 2019 "
and
"Global friction surface enumerating land-based travel walking-only speed without access to motorized transport for a nominal year 2019 "
*NOTE: Strangely, title had to be copied exactly from viewing dataframe in new tab (View(available_rasters)) and found that a space was needed at the end of title, otherwise getRaster function would not recognise title... Now realise this is called 'trailing spaces'*


Generate friction surface.
```{r generate_friction_surface}

# Generate friction surface
friction_surface <- 
  malariaAtlas::getRaster(surface = "Global friction surface enumerating land-based travel walking-only speed without access to motorized transport for a nominal year 2019 ",
                          shp = aoi_polygon)

# Note: Strangely, title had to be copied exactly from viewing dataframe in new tab (View(available_rasters)) and found that a space was needed at the end of title, otherwise getRaster function would not recognise title... Now realise this is called 'trailing spaces'

# Adjust extent
extent(friction_surface) <- extent(leastcost_motor) 
extent(friction_surface) == extent(friction_surface)
# TRUE

# Check
plot(friction_surface)

```


Create transition matrix.
```{r transition_matrix}

# Create transition matrix
transition_matrix <- gdistance::transition(friction_surface, function(x) 1/mean(x), 8) 
transition_matrix <- gdistance::geoCorrection(transition_matrix) 

```


Read in health facility data.
```{r health_facility_data}


## Data: https://zenodo.org/record/3871224#.YCuhSC2cYlk
# Downloaded Malawi health facility data and stored within folder "malawi_health_facility_data" (within working directory)

# Read in Malawi health facility data
# install.packages("readxl")
library(readxl)
mwi_mfl <- read_excel("./malawi_health_facility_data/malawi_june_2020.xlsx")
# Check
# View(mwi_mfl)

# Remove 'Non-functional' facilities
mwi_mfl <- subset(mwi_mfl, STATUS!= "Non-functional")
# Check
# View(mwi_mfl)
class(mwi_mfl)

# Convert to dataframe
mwi_mfl <- as.data.frame(mwi_mfl)
# Check
class(mwi_mfl)
# data.frame
# Check
# View(mwi_mfl)

# Change class of LATITUDE and LONGITUDE columns (currently 'character') to 'numeric'
class(mwi_mfl$LATITUDE)
class(mwi_mfl$LONGITUDE)
# Latitude and Longitude are character values: convert them to numeric
mwi_mfl$LATITUDE <- as.numeric(as.character(mwi_mfl$LATITUDE))
mwi_mfl$LONGITUDE <- as.numeric(as.character(mwi_mfl$LONGITUDE))
# Check
class(mwi_mfl$LATITUDE)
class(mwi_mfl$LONGITUDE)
# Both numeric
class(mwi_mfl)
# data.frame
# View(mwi_mfl)

# Remove any rows with 'NA' data within either LATITUDE or LONGITUDE columns
# library(tidyr)
mwi_mfl <- drop_na(mwi_mfl, "LATITUDE", "LONGITUDE")
# Check
# View(mwi_mfl)
# Checked for 'NA' and no NA cells found, also equal number of rows
class(mwi_mfl)
# data.frame


## Create SpatialPointsDataFrame (with CRS of 'friction_surface_motor')
mwi_mfl_spdf <- mwi_mfl
# There's a few points which aren't in lat/long format and should be removed
erroneous_list <- c("MC010002", "BT240003", "BT240005", "BT240006", "BLK170117", 
                    "BLK170288", "BLK170372", "MC010564", "MHG190985", "MHG191025")
# Remove
mwi_mfl_spdf <- mwi_mfl_spdf[which(!(mwi_mfl_spdf$CODE %in% erroneous_list)), ]

# Convert to a spatial object
coordinates(mwi_mfl_spdf) <- ~ LONGITUDE + LATITUDE 
proj4string(mwi_mfl_spdf) <- CRS("+proj=longlat +datum=WGS84 +no_defs ")

# Crop to extent of 'friction_surface_motor' raster
mwi_mfl_cropped <- raster::crop(mwi_mfl_spdf, y = extent(friction_surface_motor))

# Convert to simple features
mwi_mfl_cropped_sf <- st_as_sf(mwi_mfl_cropped)


```


Calculate cumulative cost.
```{r cumulative_cost}

# Calculate the cumulative cost
malaria_atlas_project_lcm <-  accCost(transition_matrix, as_Spatial(mwi_mfl_cropped))

writeRaster(malaria_atlas_project_lcm,
            "./malaria_atlas_project_outputs/malaria_atlas_project_lcm",
            format = "GTiff", 
            overwrite=TRUE)

```


Plot.
```{r, warning=FALSE}

## warning for removing NA pixels is masked

# Create dataframe
malaria_atlas_project_lcm_df <- as.data.frame(malaria_atlas_project_lcm, xy = TRUE)

# Check
# View(malaria_atlas_project_lcm_df)
# 'layers' column appears to already by in minutes 

# Rename 'layers' column to 'mins'
names(malaria_atlas_project_lcm_df)[3] <- "mins"

# Check
# View(malaria_atlas_project_lcm_df)

# Plot to visualise
ggplot()+
    geom_raster(data=malaria_atlas_project_lcm_df, aes(x = x, y = y, fill = cut(mins, c(0,30,60,120,180,240,300,max(mins)))))+
    scale_fill_brewer(palette = "YlGnBu")+
    geom_sf(data=q$osm_lines, colour="darkgrey", alpha=0.3)+
    geom_sf(data=healthfac, size=2, colour="red")+
    guides(fill=guide_legend(title="Time (mins)"))

```


Download population data (.tif) for country and covariate of interest (Malawi / entire population in 2020) from the WorldPop FTP and create raster from downloaded data
```{r download_population_data_and_create_raster}

# Obtain ISO3 country code for country of interest (coi) using 'wpgpListCountries' function
ISO3_df <- wpgpListCountries()
  
# Identify ISO3 country code for country of interest (coi) and store as object
ISO3 <- ISO3_df[ISO3_df$Country == coi, "ISO3"]

# Download  and read in dataset (.tif) for country and covariate of interest (based on ISO3 and covariate name set during setup stages)
pop_data <- 
  
  wpgpGetCountryDataset(ISO3 = ISO3,
                        covariate = covariate, 
                        destDir = "./WorldPop")

# Note: Downloaded .tif is ~50 MB: This is the entire country of Malawi

# Can just bbox be downloaded?

# Create raster
pop_data <- raster(pop_data)

# Check
pop_data
# CRS assigned: WGS84

```


Resample 'pop_data' raster to match resolution of 'leastcost_motor'.
```{r resample_pop_data}

# Determine 'pop_data' resolution 
res(pop_data)
# 0.0008333333 0.0008333333

# Determine 'leastcost_motor' resolution
leastcost_motor
res(malaria_atlas_project_lcm)
# 0.0002694946 0.0002694946

## Use 'resample' function to:
# * Resample 'pop_data' to match resolution of 'leastcost_motor'
# * Clip 'pop_data' to extent of 'leastcost_motor'
malaria_atlas_project_pop_data <- resample(pop_data, malaria_atlas_project_lcm, method = "bilinear")

# Check
res(malaria_atlas_project_pop_data) == res(malaria_atlas_project_lcm)
# True
extent(malaria_atlas_project_pop_data) == extent(malaria_atlas_project_lcm)
# True

# Check CRS
crs(malaria_atlas_project_pop_data)
# WGS84

# Visualise raster
# plot(pop_data)

```


Reclassify 'leastcost_motor' raster to chosen time-boundary categorical zones.
```{r reclassify}

## Create new raster from 'malaria_atlas_project_lcm_df' dataframe
# 'Rename 'layers' column

names(malaria_atlas_project_lcm_df)[3] <- "mins"
# Check
View(malaria_atlas_project_lcm_df)

# Create new raster from 'lcm_df' dataframe
malaria_atlas_project_lcm_df_raster <- rasterFromXYZ(malaria_atlas_project_lcm_df)
# Check
# plot(malaria_atlas_project_lcm_df_raster)

## Create matrix of time-boundary categories of interest (to resample 'lcm_raster')
## Categories: 
# * < 30 minutes
# * 30 minutes - 1 hour
# * 1 hour - 3 hours
# * 3 hours - 6 hours
# * 6 hours - 12 hours
# * > 12 hours
rcl_matrix <- c(0, 30, 1, # 30 minutes
                30, 60, 2, # 1 hour
                60, 180, 3, # 3 hours
                180, 360, 4, # 6 hours 
                360, 720, 5, # 12 hours
                720, 100000, 6 # 12+ hours
)

rcl_matrix <- matrix(rcl_matrix, ncol=3, byrow=TRUE)
# Check
# View(rcl_matrix)

## Reclassify 'malaria_atlas_project_lcm_df_raster' using 'rcl_matrix'  according to time-boundary categorical zones
malaria_atlas_project_lcm_df_raster_rcl <- reclassify(malaria_atlas_project_lcm_df_raster, rcl_matrix, include.lowest=TRUE)
# Check
malaria_atlas_project_lcm_df_raster_rcl
# Rasterlayer
# CRS not assigned

## Assign CRS to that of 'leastcost_motor' (WGS84)
projection(malaria_atlas_project_lcm_df_raster_rcl) <- crs(leastcost_motor)
# Check
crs(malaria_atlas_project_lcm_df_raster_rcl)
# CRS set (WGS84)

# Check
# plot(lcm_pop_data_rcl)

# Check as data.frame
malaria_atlas_project_lcm_df_raster_rcl_df <- as.data.frame(malaria_atlas_project_lcm_df_raster_rcl, xy = TRUE)
# View(lcm_rcl_pop_data_df)

```


Determine population within each time-boundary zone. Summate the data to give total number of individuals within chosen time-boundary categorical zones and calculate % of total population within chosen time-boundary categorical zones.
*Create for-loops?*
```{r determine_population_within_time-boundary_zones}

# Determine population (within 'pop_data' raster), within time-boundary  zones (within 'malaria_atlas_project_lcm_df_raster_rcl') using 'zonal' function 
malaria_atlas_project_zone <- zonal(malaria_atlas_project_pop_data, malaria_atlas_project_lcm_df_raster_rcl, fun = sum)

# Create dataframe
malaria_atlas_project_zone_df <- as.data.frame(malaria_atlas_project_zone, xy = TRUE)

# Check
View(malaria_atlas_project_zone_df)

# Rename columns 
colnames(malaria_atlas_project_zone_df)
names(malaria_atlas_project_zone_df)[1] <- "Zone"
names(malaria_atlas_project_zone_df)[2] <- "Zone Population"
View(malaria_atlas_project_zone_df)

# Replace time-boundary zone codes with chosen time-boundary categories
malaria_atlas_project_zone_df$Zone <- c("< 30 minutes",
                                        "30 minutes - 1 hour",
                                        "1 hour - 3 hours",
                                        "3 hours - 6 hours",
                                        "6 hours - 12 hours")

# Add 'Total Population'  column
malaria_atlas_project_zone_df$" Total Population" <-
  
  c(sum(malaria_atlas_project_zone_df$"Zone Population"[1]),
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:2]),
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:3]),
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:4]),
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]))

# Add % of total population column
malaria_atlas_project_zone_df$"% Population" <-
   
  c(sum(malaria_atlas_project_zone_df$"Zone Population"[1])   / sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) * 100,
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:2]) / sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) * 100,
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:3]) / sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) * 100,
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:4]) / sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) * 100,
    sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) / sum(malaria_atlas_project_zone_df$"Zone Population"[1:5]) * 100)
  

# Check
View(malaria_atlas_project_zone_df)
print(malaria_atlas_project_zone_df)


```




